---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# countgroups

<!-- badges: start -->
<!-- badges: end -->

The goal of countgroups is to count observations based on grouped or filtered variables. It provides a function to make this task easier. 

The package contains a function that summarises the number of observations for one or more variables within a data frame. It can be used to compute the number of  observations of one or more grouped variables, the number of observations that meet a specified condition, or a combination of the two.

## Installation

You can install the development version of countgroups from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("stat545ubc-2023/countgroups", ref = "0.1.0")
```

## Example

The following is an example of how to use the countgroups function with the palmerpenguins dataset *penguins*, to count the number of observations by island and sex for penguins of the species Adelie. 

```{r example}
library(countgroups)

count_by_groups_filter_by_condition(df=palmerpenguins::penguins, c(island, sex), cond=species=="Adelie")

```

Notice that the output is a data frame with 3 columns (island, sex, and n for number of observations), and 8 rows corresponding to each observation.

## Example 2

The function count_by_groups_filter_by_condition can be used to count observations based on grouping variables or a condition. This example demonstrates use of the function with the same *penguins* data set to filter by sex to return the number of observations of female penguins. 

```{r example 2}

count_by_groups_filter_by_condition(df=palmerpenguins::penguins, cond=sex=="female")
```

Notice that the output in this case will always be a 1x1 data frame that has one column, n, and one row, the number of observations that meet the condition.
